Music Perception 2: Preregistered Analysis

### 2022-03-08 experiment-1/analysis/prereg_analysis/prereg_analysis.Rmd
update prereg analysis (add two-way anova / jittered plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

# Experiment
## Preregistered Analysis
We originally had a analysis plan to look at the accuracy of participants' judgements whether participants would be able to judge stimuli from the teaching condition as "teaching". For more detailed explanations, please see [the OSF preregistration](https://osf.io/z3j69).

@@

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques (expressions) separately. The order of the blocks was counterbalanced across participants.
- 48 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented across participants.

@@

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# stats
if (!require("ppcor")) {install.packages("ppcor"); require("ppcor")}
if (!require("performance")) {install.packages("performance"); require("performance")}
```

@@

## 1. Statistics
### All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

# all
subset <- ind_data[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/96)*100, SD_Percent = (sd(Sum)/96)*100)]
subset
```

```{r plot, echo = FALSE}
ggboxplot(ind_data, x = "N", y = "Percent", add = "dotplot", xlab = FALSE) + ylab("Correct %") + theme(axis.text.x = element_blank())
```

@@

### Articulation vs. Dynamics

```{r plot2, echo = FALSE}
ind_data_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Skill)]
ind_data_ad <- ind_data_ad[order(subjectNumber, Skill)]

subset_ad <- ind_data_ad[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(Skill)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad[Skill == "articulation"]$subjectNumber, articulation = ind_data_ad[Skill == "articulation"]$Percent, dynamics = ind_data_ad[Skill == "dynamics"]$Percent)
box <- ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box
```

## One-sample *t*-test
### Articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "articulation"]$Percent)

@@

## Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent-ind_data_ad[Skill == "articulation"]$Percent)

@@

>>>>

# Experiment
## Preregistered Analysis
The preregistered analysis aims to investigate whether participants could distinguish performances produced for teaching purposes from those produced for performing purposes. For more detailed explanations, please see [the OSF preregistration](https://osf.io/z3j69).

@@

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques separately. The order of the blocks was counterbalanced across participants.
- 48 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented within a block.

@@

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# stats
if (!require("afex")) {install.packages("afex"); require("afex")}
```

@@

## 1. Statistics
## All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

# all
subset <- ind_data[, .(N = .N, Mean_Percent = mean(Percent), SD_Percent = sd(Percent))]
subset
```

```{r plot, echo = FALSE}
ggboxplot(ind_data, x = "N", y = "Percent", add = "dotplot", xlab = FALSE) + ylab("Correct %") + theme(axis.text.x = element_blank()) + geom_hline(yintercept = 50, linetype = "dashed")
```

@@

## Articulation vs. Dynamics
Reference for a paired plot with jitter: [https://stackoverflow.com/questions/48042810/add-jitter-to-ggpaired#48044022](https://stackoverflow.com/questions/48042810/add-jitter-to-ggpaired#48044022)

```{r plot2, echo = FALSE}
ind_data_ad <- data[subjectNumber != 4, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Skill)]
ind_data_ad <- ind_data_ad[order(subjectNumber, Skill)]

subset_ad <- ind_data_ad[, .(N = .N, Mean_Percent = mean(Percent), SD_Percent = sd(Percent)), by = .(Skill)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad[Skill == "articulation"]$subjectNumber, articulation = ind_data_ad[Skill == "articulation"]$Percent, dynamics = ind_data_ad[Skill == "dynamics"]$Percent)

# plot with jitter
n <- length(data_box$articulation)
d <- data.frame(y = c(data_box$articulation, data_box$dynamics), 
                x = rep(c(1, 2), each = n),
                id = factor(rep(1:n, 2)))

set.seed(321)    
d$xj <- jitter(d$x, amount = .05)
ggplot(data = d, aes(y=y)) +
  geom_boxplot(aes(x = x, group = x), width = 0.5, outlier.shape = NA) +
  geom_point(aes(x = xj)) +
  geom_line(aes(x = xj, group = id), color = "gray", size = .4) + labs(x = "Skill", y = "Correct %") +
  scale_x_continuous(breaks = c(1, 2), labels = c("articulation", "dynamics"), limits = c(0.5, 2.5)) + geom_hline(yintercept = 50, linetype = "dashed") + theme_pubr()

# plot without jitter
# ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, xlab = "Skill", ylab = "Correct %") + geom_hline(yintercept = 50, linetype = "dashed")
```

### One-sample *t*-test
### Articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "articulation"]$Percent)

@@

### Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent-ind_data_ad[Skill == "articulation"]$Percent)

@@

## Two-way ANOVA (Condition * Skill)
```{r plot3, echo = FALSE}
ind_data_tp_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
ind_data_tp_ad <- ind_data_tp_ad[order(subjectNumber, Condition, Skill)]

ggboxplot(ind_data_tp_ad, "Skill", "Percent", color = "Condition") + geom_hline(yintercept = 50, linetype = "dashed")
```

```{r, echo = FALSE}
anova <- aov_ez("subjectNumber", "Percent",
                data = ind_data_tp_ad, within = c("Condition", "Skill"))
anova
```

### 2022-03-08 experiment-2/analysis/prereg_analysis/prereg_analysis.Rmd
update prereg analysis (add two-way anova / jittered plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

- 21 participants (experimental error = 1) >> 20 participants were included for analysis.

# Experiment
## Preregistered Analysis
The preregistered analysis aims to investigate whether participants could distinguish performances produced for teaching purposes from those produced for performing purposes. For more detailed explanations, please see [the OSF preregistration](https://osf.io/z3j69).

## Stimuli
- Stimuli were selected from the previous study ([the sound of teaching music/experiment-1](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-1)).
- 48 stimuli were randomly chosen from each technique (either articulation or dynamics - therefore in total 96 stimuli). Half of them was selected from the teaching condition and the other half was from the performing condition.
  + 24 for teaching-articulation performances
  + 24 for performing-articulation performances
  + 24 for teaching-dynamics performances
  + 24 for performing-dynamics performances

@@

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques separately. The order of the blocks was counterbalanced across participants.
- 48 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented within a block.

@@

## 1. Statistics
## All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

@@

## Two-way ANOVA (Condition * Skill)
```{r plot3, echo = FALSE}
ind_data_tp_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
ind_data_tp_ad <- ind_data_tp_ad[order(subjectNumber, Condition, Skill)]

>>>>

- 21 participants (excluded = 1) >> 20 participants were included for analysis.

# Experiment
## Preregistered Analysis
We originally had a analysis plan to look at the accuracy of participants' judgements whether participants would be able to judge stimuli from the teaching condition as "teaching". For more detailed explanations, please see [the OSF preregistration](https://osf.io/z3j69).

## Stimuli
- Stimuli were selected from the previous study ([the sound of teaching music/experiment-2](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2)).
- 36 stimuli were randomly chosen from each technique (either articulation or dynamics - therefore in total 72 stimuli). The half of them was selected from the teaching condition and the other half was from the performing condition.
  + 18 for teaching-articulation performances
  + 18 for performing-articulation performances
  + 18 for teaching-dynamics performances
  + 18 for performing-dynamics performances

@@

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques (expressions) separately. The order of the blocks was counterbalanced across participants.
- 36 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented across participants.

@@

## 1. Statistics
## All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 4, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

@@

## Two-way ANOVA (Condition * Skill)
```{r plot3, echo = FALSE}
ind_data_tp_ad <- data[subjectNumber != 4, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
ind_data_tp_ad <- ind_data_tp_ad[order(subjectNumber, Condition, Skill)]
